Parsing Postfix log files

### 2009-04-06 doc/logparser-implementation.tex
Dealt with some XXX items. Improvements in results section. Added table showing number of rules per program.

    \item The regex is valid and compiling it for efficiency
        (\sectionref{Caching compiled regexes}).

@@

XXX CARL SUGGESTS ADDING GRAPHS AND STATS TABLES FOR THESE\@.

@@

    \item How long it takes to parse each log file.

    \item The number of rules tried when recognising log lines; the
        framework may need to try multiple rules for a given log line
        before it finds one that recognises it.

    \item How many log lines did the framework attempt to recognise, how
        many were skipped because there were no rules whose condition was
        equal to that log line's program, how many were successfully
        recognised, and how many were not recognised.

@@

    \item XXX FURTHER COMPLICATIONS MAY ARISE\@.

    \item XXX DATABASE GROWS WITHOUT BOUNDS\@.

>>>>

    \item That the regex is valid.  The regex will also be compiled for
        efficiency (\sectionref{Caching compiled regexes}).

@@

XXX ASK CARL IF I SHOULD HAVE MEAN, STANDARD DEVIATION, MORE GRAPHS\@?

@@

    \item How long it takes to parse each log file (\graphref{parsing time
        vs file size vs number of log lines graph}).

    \item The number of rules tried when recognising log lines
        (\graphref{Average number of rules tried per log line}); the
        framework may need to try multiple rules for a given log line
        before it finds one that recognises it.

    \item The number of log lines and rules tried for each Postfix
        component (\graphref{Average number of rules tried per log line for
        each Postfix component}).

@@

    \item Further complications may arise when parsing log files, and
        \parsername{} will need to be modified to deal with them.

    \item \parsername{} does not limit the size of the database, which will
        grow without bounds unless the user deletes connections and results
        from it.  This is both a limitation and a benefit: the benefit is
        that data will never be unexpectedly deleted, but the limitation is
        that the user must manage the size of the database.

### 2009-04-06 doc/logparser-results.tex
Dealt with some XXX items. Improvements in results section. Added table showing number of rules per program.

Parsing efficiency is an obvious concern when the parser routinely needs to
parse large log files.  The server that generated the log files used in
testing this parser accepts approximately 10,000 mails for 700 users a day;
\graphref{Mails received per day} and \tableref{Number of mails received
per day: statistics} show that, as expected, far more mails are received on
weekdays than at weekends.  The mail server in question is a production
mail server handling mail for a university department; the benefit of using
this server is that its log files exhibit the idiosyncrasies and
peculiarities a mail server in the wild must deal with, but the downside is
that significantly altering the configuration to accommodate this project
is not an option.  Median log file size is 50MB, containing 285,000 log
lines; large scale mail servers would have much larger log files.  Note
that \graphref{Mails received per day} and \tableref{Number of mails
received per day: statistics} show the number of mails received by
\acronym{SMTP} only; in particular the mail loops noticeable in later
graphs do not contribute to these figures.

@@

When generating the timing data used in this section, \numberOFlogFILES{}
log files (totaling 10.08 GB, \numberOFlogLINEShuman{} log lines) were each
parsed 10 times, and the times from the 10 runs averaged.  The computer
used for test runs was a Dell Optiplex 745, shown in \tableref{Details of
the computer used to generate statistics}; it was dedicated to the task of
gathering statistics from test runs, and was not used for any other purpose
while test runs were ongoing.  Saving results to the database was disabled
for the test runs, because that dominates the run time of the parser, and
the tests are aimed at measuring the speed of \parsername{} rather than the
speed of the database and the disks the database is stored on.  Parsing all
\numberOFlogFILES{} log files in one run without saving results to the
database took \input{build/include-full-run-duration}, mean throughput is
\input{build/include-full-run-throughput}; median throughput when parsing
files separately was \input{build/include-median-throughput-MB}
(\input{build/include-median-throughput-log-lines}) parsed per minute.  In
contrast, when saving results to the database, parsing all
\numberOFlogFILES{} log files took
\input{build/include-insert-results-duration}, with mean throughput of
\input{build/include-insert-results-throughput}; parsing makes up
\input{build/include-skip-inserting-as-percentage-of-inserting} of the
execution time when saving results to the database, so a 50\% improvement
in the speed of parsing would result in only a
\input{build/include-skip-inserting-with-50-percent-improvement--as-percentage-of-inserting}
improvement in speed overall.

@@

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{parsing time vs file size vs number of log lines graph} shows the
parsing time in seconds, file size in MB, and number of log lines in tens
of thousands, for each of the \numberOFlogFILES{} log files.  The three
lines run roughly in parallel, giving the impression that the algorithm
scales linearly with input size.  This impression is borne out by
\graphref{parsing time vs file size vs number of log lines factor}, which
plots both the ratio of file size vs parsing time, and the ratio of number
of log lines vs parsing time (higher is better in both cases);
\tableref{parsing time vs file size vs number of log lines factor table}
shows the same ratios for different groups of log files.  The ratios are
quite tightly banded, showing that the algorithm scales linearly; the
ratios increase (i.e.\ improve) for log files 22 and 62--68, despite their
larger than average size.  Both groups of log files are much larger than
usual due to a mail loop caused by a user who set up mail forwarding
incorrectly, resulting in a very different distribution of log lines:
normally most log lines are generated by mail delivery attempts from other
hosts, but when the mail loops occurred most of the log lines resulted from
failed delivery of mail generated on the server itself.  The Postfix
components that generated most of the log lines during the mail loop have
fewer associated rules than the Postfix components whose log lines normally
make up the bulk of each log file, so the average number of rules tried per
log line reduces and so does the average parsing time per log line.
\Graphref{Average number of rules tried per log line} shows the drop in the
average number of rules tried per log line for log files containing a mail
loop, and \graphref{Average number of rules tried per log line for each
Postfix component} shows the average number of rules tried per log line for
each Postfix component.

@@

@@

At the time of writing there are \numberOFrules{} different rules: the top
10\% recognise \input{build/include-top-ten-hits}\% of the log lines, with
the remaining log lines split across the other 90\% of the rules (as shown
in \graphref{rule hits graph}).  Assuming that the distribution of log
lines is reasonably consistent over time, \parsernames{} efficiency should
benefit from trying more frequently recognising rules before those that
recognise less frequently.  To test this hypothesis three full test runs
were performed with different rule orderings:

@@

\newpage{} % XXX GRAPH PLACEMENT FLIP-FLOPPING

\Graphref{Parsing time relative to shuffled ordering graph} shows the
parsing times of optimal and reverse orderings relative to shuffled
ordering; the mean relative parsing times for different groupings of log
files are given in \tableref{Parsing time relative to shuffled ordering
table}.  This optimisation provides a mean reduction in parsing time of
\input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
with normal log files,
\input{build/include-optimal-ordering-parsing-time-reduction-logs-22-62-68}\%
when a mail loop occurs and the distribution of log lines is unusual.
Optimal rule ordering has other benefits, described in
\sectionref{scalability as the number of rules rises}.

@@

\newpage{} % XXX VARIOREF PROBLEM

Differences in rule ordering have less effect on parsing time when parsing
log files 22 and 62--68, due to the different distribution of log lines in
those log files.
When a mail loop occurs the vast majority of log lines are from Postfix
components that have few rules associated with them, whereas in general
most log lines are from Postfix components that have a high number of
associated rules; this results in the average number of rules required to
parse a log line from log files 22 or 62--68 being considerably lower than
usual.  The ratio of number of log lines vs.\ parsing time in
\graphref{parsing time vs file size vs number of log lines factor} shows a
noticeable improvement at log files 22 and 62--68, where a higher ratio
means more log lines parsed per unit of time; similarly \graphref{Parsing
time relative to shuffled ordering graph} shows a much smaller difference
between the orderings at log files 22 and 62--68.

>>>>

Parsing efficiency is an obvious concern when the parser routinely needs to
parse large log files.  The server that generated the log files used in
testing this parser accepts approximately 10,000 mails for 700 users per
day; \graphref{Mails received per day} and \tableref{Number of mails
received per day: statistics} show that, as expected, far more mails are
received on weekdays than at weekends.  The mail server in question is a
production mail server handling mail for a university department; the
benefit of using this server is that its log files exhibit the
idiosyncrasies and peculiarities a mail server in the wild must deal with,
but the downside is that significantly altering the configuration to
accommodate this project is not an option.  Median log file size is 50MB,
containing 285,000 log lines; large scale mail servers would have much
larger log files.  Note that \graphref{Mails received per day} and
\tableref{Number of mails received per day: statistics} show the number of
mails received by \acronym{SMTP} only, and the mail loops noticeable in
later graphs do not contribute to these figures.

@@

When generating the timing data used in this section, \numberOFlogFILES{}
log files (totaling 10.08 GB, \numberOFlogLINEShuman{} log lines) were each
parsed 10 times, and the times from the 10 runs averaged.  The computer
used for test runs was a Dell Optiplex 745, shown in \tableref{Details of
the computer used to generate statistics}; it was dedicated to the task of
gathering statistics from test runs, and was not used for any other purpose
while test runs were ongoing.  Saving results to the database was disabled
for the test runs, because that dominates the run time of the parser, and
the tests are aimed at measuring the speed of \parsername{} rather than the
speed of the database and the disks the database is stored on.  Parsing all
\numberOFlogFILES{} log files in one run without saving results to the
database took \input{build/include-full-run-duration}, mean throughput was
\input{build/include-full-run-throughput}; median throughput when parsing
files separately was \input{build/include-median-throughput-MB}
(\input{build/include-median-throughput-log-lines}) parsed per minute.  In
contrast, when saving results to the database, parsing all
\numberOFlogFILES{} log files took
\input{build/include-insert-results-duration}, with mean throughput of
\input{build/include-insert-results-throughput}.  Parsing makes up only
\input{build/include-skip-inserting-as-percentage-of-inserting} of the
execution time when saving results to the database, so a 50\% improvement
in the speed of parsing would result in only a
\input{build/include-skip-inserting-with-50-percent-improvement--as-percentage-of-inserting}
improvement in speed overall.

@@

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{parsing time vs file size vs number of log lines graph} shows the
parsing time in seconds, file size in MB, and number of log lines in tens
of thousands, for each of the \numberOFlogFILES{} log files.  The three
lines run roughly in parallel, giving the impression that the algorithm
scales linearly with input size.  This impression is borne out by
\graphref{parsing time vs file size vs number of log lines factor}, which
plots both the ratio of file size vs parsing time, and the ratio of number
of log lines vs parsing time (higher is better in both cases);
\tableref{parsing time vs file size vs number of log lines factor table}
shows the same ratios for different groups of log files.  The ratios are
quite tightly banded, showing that the algorithm scales linearly; the
ratios increase (i.e.\ improve) for log files 22 and 62--68, despite their
larger than average size.  Both groups of log files are much larger than
usual due to a mail loop caused by a user who set up mail forwarding
incorrectly, resulting in a very different distribution of log lines:
normally most log lines are generated by mail delivery attempts from other
hosts, but when the mail loops occurred most of the log lines resulted from
failed delivery of mail generated on the server itself.  The Postfix
components that generated most of the log lines during the mail loop have
fewer associated rules than the Postfix components whose log lines normally
make up the bulk of each log file, so the average number of rules tried per
log line reduces and so does the average parsing time per log line.
\Graphref{Number of rules per Postfix component} shows the number of rules
per Postfix component, \graphref{Average number of rules tried per log
line} shows the drop in the average number of rules tried per log line for
log files containing a mail loop, and \graphref{Average number of rules
tried per log line for each Postfix component} shows the average number of
rules tried per log line for each Postfix component.

@@

\begin{table}[htbp]
    \caption{Number of rules per Postfix component}
    \empty{}\label{Number of rules per Postfix component}
    \centering{}
    \begin{tabular}{lr}
        \tabletopline{}%
        Postfix component & Number of rules \\
        \tablemiddleline{}%
        \input{build/include-number-of-rules-per-program}
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

\parsername{} has \numberOFrules{} rules: the top 10\% recognise
\input{build/include-top-ten-hits}\% of the log lines, with the remaining
log lines split across the other 90\% of the rules (as shown in
\graphref{rule hits graph}).  Assuming that the distribution of log lines
is reasonably consistent over time, \parsernames{} efficiency should
benefit from trying rules that recognise log lines more frequently before
those rules that recognise log lines less frequently.  To test this
hypothesis, three full test runs were performed with different rule
orderings:

@@

\Graphref{Parsing time relative to shuffled ordering graph} shows the
parsing times of optimal and reverse orderings relative to shuffled
ordering; the mean relative parsing times for different groupings of log
files are given in \tableref{Parsing time relative to shuffled ordering
table}.  This optimisation provides a mean reduction in parsing time of
\input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
with normal log files,
\input{build/include-optimal-ordering-parsing-time-reduction-logs-22-62-68}\%
when a mail loop occurs and the distribution of log lines is unusual.
Optimal rule ordering has other benefits, described in
\sectionref{scalability as the number of rules rises}.  Differences in rule
ordering have less effect on parsing time when parsing log files 22 and
62--68, due to the different distribution of log lines in those log files.

@@
